<script src = "../../src/linked-charts.min.js"></script>
<link rel="stylesheet" type="text/css" href="../../src/linked-charts.css">

## R/LinkedCharts Tutorial
# Customising your chart

Here we will show how one can use various built-in properties to customise charts.
The main goal of this tutorial is to give overview of the adjuststable apects in R/LinkedCharts, 
(colours, axes, labels, etc.). Therefore we are going to use well known example data sets
such ss [Iris](https://en.wikipedia.org/wiki/Iris_flower_data_set) flower data set or randomly 
generated data. 

```{r}
data("iris") #load Iris data set
head(iris)
```

We assume that you are already familiar with R/LinkedCharts and its main ideas and principles
and want to explore more possibilities of the library. Otherwise, it can be better first to 
go through [this](../oscc/oscc.html) tutorial.

```{r message=FALSE}
library(rlc) #load the library
```

## Colour

The simplest way to colour elements of a chart is to use the `colourValue` property.
It takes numbers or strings for each point (line, bar, etc.) and based on that generates
a continuous or categorical colour scale.

```{r}
openPage(layout = "table1x2")

# a scatter plot with a categorical colour scale
lc_scatter(dat(x = iris$Sepal.Length,
               y = iris$Petal.Length,
               colourValue = iris$Species),
           width = 300, height = 300, #change width and height of the chart to 300px
           place = "A1")

# a scatter plot with a continuous colour scale
lc_scatter(dat(x = iris$Sepal.Length,
               y = iris$Petal.Length,
               colourValue = iris$Petal.Width),
           width = 300, height = 300,
           place = "A2")
```
To further specify colours one can use `palette` and `colourDomain` properties. 

In case of categorical colour scale `colourDomain` is a set of all possible colour values.
Thus for the chart on the left by default `colourDomain = c("setosa", "versicolor", "virginica")`.
If a value is not included in colourDomain, corresponding points will be black.

For continuous colour scales, `colourDomain` defines the range in wich colour values can change. All values
outside this range will produce colours that correspond to maximal or to minimal value of the `colourDomain`.

To illustrate all this, let's add `colourDomain` to our example.

```{r}
openPage(layout = "table1x2")

# a scatter plot with a categorical colour scale
lc_scatter(dat(x = iris$Sepal.Length,
               y = iris$Petal.Length,
               colourValue = iris$Species,
               colourDomain = c("setosa", "something else", "virginica")),
           width = 300, height = 300, #change width and height of the chart to 300px
           place = "A1")

# a scatter plot with a continuous colour scale
lc_scatter(dat(x = iris$Sepal.Length,
               y = iris$Petal.Length,
               colourValue = iris$Petal.Width,
               colourDomain = c(-1, 1)),
           width = 300, height = 300,
           place = "A2")
```

On the left plot, all the "versicolor" points are now black, since this value is no longer present in
`colourDomain`. At the same time, orange colour is reserved for "something else". On the plot to the right,
colour scale now varies between -1 and 1. All the points with `colourValue` greater than 1 are just purple.

Finally, `palette` defines what colours are used. It is always a vector of colours (their 
[names](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf) or
[hexadecimal](https://en.wikipedia.org/wiki/Web_colors#HTML_color_names) codes). For categorical colour scale,
`palette` must have a colour for each element from `colourDomain`. For continous scales, `palette` is a set of
"reference points" for the colour scale. By default, they are spread evenly withing `colourDomain`, but one can
also specify intermediate points.

```{r}
openPage(layout = "table1x2")

# a scatter plot with a categorical colour scale
lc_scatter(dat(x = iris$Sepal.Length,
               y = iris$Petal.Length,
               colourValue = iris$Species,
               colourDomain = c("setosa", "something else", "virginica", "versicolor")),
               palette = c("gold", "hotpink", "dodgerblue"),
           width = 300, height = 300, #change width and height of the chart to 300px
           place = "A1")

# a scatter plot with a continuous colour scale
lc_scatter(dat(x = iris$Sepal.Length,
               y = iris$Petal.Length,
               colourValue = iris$Petal.Width,
               colourDomain = c(0, 0.3, 2.6),
               palette = c("red", "grey", "black")),
           width = 300, height = 300,
           place = "A2")
```
Now chart to the left has more elements in `colourDomain` than colours in `palette` and therefore colours are repeated, 
starting from the first element in `palette`. For the chart to the right we've provided `palette` with three colours
and also we've added an intermediate point in `colourDomain`. 



```{r}
lc_scatter(dat(x = 1:5,  # x-coordinates of the points are 1, 2, 3, 4, 5
               y = rep(1, 5), # all 5 points has 1 as their y-coordinates
               size = 15, # Let's have big points! (Default point size is 6)
               width = 350, # Our chart will be 350px wide...
               height = 200, # ...and 200px high.
               colour = c("red", "#123456", rgb(0.4, 0.8, 0.1), "#aaa", "some_strange_colour"))
          )
```

Here, we are interested in `colour` property. 

## Shape

## Labels

## Chart Setting

## Others
